The role of aneuploidy in the evolution of cancer drug resistance

### 2023-06-20 EvolutionaryRescue1.tex
fixed typos

We follow the number of cancer cells that have one of three different genotypes at time $t$: wildtype, $w_t$; aneuploid, $a_t$; and mutant, $m_t$. 
These cells divide and die with rates $\lambda_k$ and $\mu_k$ (for $k=w, a, m$).
The difference between the division and death rate is $\Delta_k = \lambda_k-\mu_k$.
We assume the population of cells is under a strong stress, such as drug therapy, to which the wildtype genotype is susceptible and therefore $\Delta_w<0$, whereas the mutant is resistant to the stress, $\Delta_m>0$
We analyze three scenarios: in the first, aneuploid cells are partially resistant, $\Delta_m>\Delta_a>0$; in the second, aneuploid cells are tolerant, $0>\Delta_a>\Delta_w$ \citep[see][for the distinction between susceptible, resistant, and tolerant]{brauner2016distinguishing}; in the third, aneuploid cells are non-growing or "barely growing", that is, either slightly tolerant or slightly resistant, such that $\Delta_a \approx 0$.
Wildtype cells may missegregate to become aneuploids at rate $u$. Both aneuploid and wildtype cells may mutate to become mutants at rate $v$ (\Cref{figureAneuploidy}).

>>>>

We follow the number of cancer cells that have one of three different genotypes at time $t$: wildtype, $w_t$; aneuploid, $a_t$; and mutant, $m_t$. 
These cells divide and die with rates $\lambda_k$ and $\mu_k$ (for $k=w, a, m$).
The difference between the division and death rate is $\Delta_k = \lambda_k-\mu_k$.
We assume the population of cells is under a strong stress, such as drug therapy, to which the wildtype genotype is susceptible and therefore $\Delta_w<0$, whereas the mutant is resistant to the stress, $\Delta_m>0$.
We analyze three scenarios: in the first, aneuploid cells are partially resistant, $\Delta_m>\Delta_a>0$; in the second, aneuploid cells are tolerant, $0>\Delta_a>\Delta_w$ \citep[see][for the distinction between susceptible, resistant, and tolerant]{brauner2016distinguishing}; in the third, aneuploid cells are non-growing or "barely growing", that is, either slightly tolerant or slightly resistant, such that $\Delta_a \approx 0$, in a sense that we will make more precise.
Wildtype cells may missegregate to become aneuploids at rate $u$. Both aneuploid and wildtype cells may mutate to become mutants at rate $v$ (\Cref{figureAneuploidy}).
